---
title: "Descriptive analysis of daily activity"
author: "HLiu"
date: "Sunday, September 20, 2015"
output: html_document
---


## Introduction

It is now possible to collect a large amount of data about personal
movement using activity monitoring devices such as a
[Fitbit](http://www.fitbit.com), [Nike
Fuelband](http://www.nike.com/us/en_us/c/nikeplus-fuelband), or
[Jawbone Up](https://jawbone.com/up). These type of devices are part of
the "quantified self" movement -- a group of enthusiasts who take
measurements about themselves regularly to improve their health, to
find patterns in their behavior, or because they are tech geeks. But
these data remain under-utilized both because the raw data are hard to
obtain and there is a lack of statistical methods and software for
processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring
device. This device collects data at 5 minute intervals through out the
day. The data consists of two months of data from an anonymous
individual collected during the months of October and November, 2012
and include the number of steps taken in 5 minute intervals each day.

## Data

The data for this assignment can be downloaded from the course web
site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as `NA`)

* **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

* **interval**: Identifier for the 5-minute interval in which
    measurement was taken




The dataset is stored in a comma-separated-value (CSV) file and there
are a total of 17,568 observations in this
dataset.


## Assignment


**1. Loading and preprocessing the data**  

```{r,echo=TRUE}
setwd("C:/LiuH/Couresa_R/repdata-data-activity")
#chol <- read.table(url("http://assets.datacamp.com/blog_assets/chol.txt"), 
#                         header = TRUE)
activity <- read.csv("activity.csv")
names(activity)
activity2<-subset(activity,activity$steps != 'NA')


```

**2. What is mean total number of steps taken per day?**  

```{r,echo=TRUE}
library(plyr)
sum<-ddply(activity2,.(date),summarize,tot=sum(steps))

library(ggplot2)

ggplot(sum, aes(x=date,y=tot,fill=date),main='Histogram of the total number of      steps taken each day') +
      geom_histogram(stat='identity',binwidth=2) +
      theme(axis.text.x = element_text(angle= -90, vjust=0.5, hjust=1, colour                           ="black"))

sum$ID = 'SID001'
ddply(sum,.(ID),summarize,'mean' = mean(tot),'median' = median(tot),'sd'=sd(tot))


```
  
  
**3. What is the average daily activity pattern?**  

```{r,echo=TRUE}
df<-ddply(activity2,.(interval),summarize,'mean' = mean(steps))


library(caTools)
activity2$stepsMMean<- runmean(activity2$steps,nrow(activity2)，align='right')

plot(df$mean,df$interal,type="l",ylab='Mean number of steps',xlab='Interval')

#Interval which contains the maximum mean number of step
df[df$mean==max(df$mean),]



```

  
  
**4. Imputing missing values**  

```{r,echo=TRUE}

##Calculate the total number of rows with 'NA'
nrow(activity[activity$steps=='NA',])

##Filling in missings with the median steps of the day

impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
activity3 <- ddply(na.omit(activity), ~ date, transform, steps = impute.median(steps))


sum2<-ddply(activity3,.(date),summarize,tot=sum(steps))

###Make a histogram of the total number of steps taken each day
ggplot(sum2, aes(x=date,y=tot,fill=date),main='Histogram of the total number of      steps taken each day') +
      geom_histogram(stat='identity',binwidth=2) +
      theme(axis.text.x = element_text(angle= -90, vjust=0.5, hjust=1, colour                           ="black"))

###The mean and median total number of steps taken per day after imputing
sum2$ID = 'SID001'
ddply(sum2,.(ID),summarize,'mean' = mean(tot),'median' = median(tot),'sd'=sd(tot))


```
  
    
**5. Are there differences in activity patterns between weekdays and weekends?**   

```{r,echo=TRUE}
activity3$dayType<- ifelse(weekdays(as.Date(activity3$date))=='Sunday'|weekdays(as.Date(activity3$date))=='Saturday','Weekend','Weekday')

activity3.we=activity3[activity3$dayType == 'Weekday',]
activity3.wd=activity3[activity3$dayType == 'Weekend',]
df.we<-ddply(activity3.we,.(interval),summarize,'mean' = mean(steps),dayType=
                 'Weekday')
df.wd<-ddply(activity3.wd,.(interval),summarize,'mean' = mean(steps),dayType=
                 'Weekend')


activity4<-rbind(df.we,df.wd)


ggplot(activity4, aes(x = interval, y = mean,group=dayType)) +
  geom_line(linetype='solid',size=0.5) +
  ylab("Mean of number of steps") +
  xlab("Interval") +
  facet_wrap( ~ dayType,ncol=1)

 

```